---
title: "Homework5"
author: "Kapil Taspa"
date: "2024-10-23"
output: html_document
---

## Question 1

***Create a data frame containing: the names of the players, the minutes played, and their three-point shooting percentage. How many players are in the roster for 2024?***

There are 183 players in the 2024 WNBA roster. Data frame displayed below with names of players as names, the minutes played as minutes, and their three-point shooting percetange as three_percentage.

```{r, warning=FALSE}
#setup necessary libraries
library(rvest)
library(ggplot2)

#create the data frame
wnba_data <- data.frame(
  'name' = read_html('https://www.basketball-reference.com/wnba/years/2024_totals.html') |>
  html_elements('th.left') |>
  html_text(trim = TRUE),
  
  'minutes' = read_html('https://www.basketball-reference.com/wnba/years/2024_totals.html') |>
  html_elements('.right:nth-child(5)') |>
  html_text(trim = TRUE) |>
  as.numeric(),
  
  'three_percentage' = read_html('https://www.basketball-reference.com/wnba/years/2024_totals.html') |>
  html_elements('.right:nth-child(14)') |>
  html_text(trim = TRUE) |>
  as.numeric()
  
)
wnba_data
wnba_data |>
  nrow()
```

## Question 2

***Create a graph of the distribution of the three-point shooting percentage made by each players. Report and interpret appropriate statistics.***

A histogram of the three-point shooting percentage made by each player in the 2024 WNBA season is displayed below. The mean is \~0.297 and the median is \~0.316 which indicates a slight left or negative skew because the mean is less than the median. Most of the three-point shooting percentages are below 0.5, however, they are more concentrated on the right side of that i.e. less values are closer to 0.

```{r, warning=FALSE}
#histogram that shows the distribution of the three-point shooting percentage made by each player in the 2024 WNBA
#added a breaks argument so the graph shows the correct skew
hist(wnba_data$three_percentage, 
     main = "Distribution of Three-Point Percentages in the 2024 WNBA",
     xlab = "Three-Point Shooting Percentage",
     ylab = "Frequency",
     col = "red", na.rm = TRUE, breaks = 50)

#calculate the relevant statistics in this case the mean and median to determine the skew
mean(wnba_data$three_percentage, na.rm = TRUE)
median(wnba_data$three_percentage, na.rm = TRUE)
```

## Question 3

***Create a graph that shows the relationship between the minutes played by each player and their three-point shooting percentage. Choose one particular player, highlight her point on the graph and label it with her name.***

A scatter plot is crated below with the player Sabrina Ionescu highlighted in red and text indicating where the point is on the graph.

```{r, warning=FALSE}
#scatter plot that shows the relationship between the minutes played by each player and their three-point shooting percentage
ggplot(wnba_data, aes(x = minutes, y = three_percentage)) +
  geom_point(color = "black") +
  geom_point(data = subset(wnba_data, name == "Sabrina Ionescu"), 
             aes(x = minutes, y = three_percentage), 
             color = "red", size = 3) +  #higlight the point
  geom_text(data = subset(wnba_data, name == "Sabrina Ionescu"), #add text with her name
            aes(label = name), vjust = -1 ,color = "red") +  # Label "Player X" with her name
  labs(title = "Relationship Between Minutes Played & Three-Point Percentage in the 2024 WNBA",
       x = "Minutes Played",
       y = "Three-Point Shooting Percentage")


```
